My Ph.D. era has ended, and I cannot be more grateful for all the lessons that I have learned.
%
It was a tough 6 years with challenges and many ups and downs. After all I have been through, deprived of studying in my home country, starting a new life abroad, and thriving during my Ph.D., it feels great and rewarding.
%
Hereby, I would like to acknowledge those people who have been part of this journey and have helped me in
many ways during my 6 years at the University of Utah.

First of all, I would like to thank my advisor, Dr. Ganesh Gopalakrishnan, who taught me to work hard, generate ideas and accomplish the work.
%
This dissertation would not be possible without his expertise, guidance, and support.
%
I have learned a ton of technical and life lessons from him.
%
Second, I would like to thank my external committee member who served as my M.S. advisor at Texas State University, Dr. Martin Burtscher.
%
He taught me how to research computer science and encouraged me to pursue my Ph.D.
%
Dr. Burtscher also supported me with his valuable feedback during my Ph.D. years.
%
Third, I want to thank my committee members, Dr. Rakamaric, Dr. Sundar, and Dr. Lex, for their support and significant feedbacks during the completion of this dissertation.

My father, Gholamhossein Taheri, passed away in January 2014. His sudden loss impacted our family badly. However, we stick together, supported each other, and survived the tough days without him. Now here I am, defended my Ph.D., my mother thrived nicely expanding my father's business and kept the job for many of her workers. And my sisters, Farnaz and Tannaz, are pursuing their higher degrees in the U.S. and Canada. I wish my dad could see us flourishing. He would be so much happy and proud. I would like to thank him for all the love and support he had for me.
I want to thank my mother, Afsar Amiri, who sacrificed her life for her kids and supported us however she could. I would not be here if it were not because of her love, support, and kindness.

During my Ph.D., I was fortunate as I met the love of my life, Zahra. Being around her and feeling her love and support, especially when I was frustrated and disappointed, was the best gift that anybody could have asked for. I want to appreciate her a lot and wish her all the best as she is also working on her Ph.D. in Mechanical Engineering at the University of Utah at the moment.

I also want to acknowledge all the teachers, professors, and advisors who taught me the science of nature and the skills to engineer new creatures. I appreciate my teachers in Alavi elementary school, Taha junior high school, Andishe high school, professors, and advisors in the Computer Science and Engineering Department at the Shiraz University, Texas State University, and the University of Utah.
%
Finally, I would like to thank all my friends and colleagues. My childhood friends in my hometown Shiraz, my friends in San Marcos, Texas, and my new friends in Salt Lake City, Utah. I also want to thank my colleagues in the Efficient Computing Laboratory (ECL) at Texas State University and the Center of Parallel Computing (CPU) at the University of Utah.
